Individual and Differential Harm in Redistricting

### 2024-12-17 paper/harm.qmd
submission

---
title: Individual and Differential Harm in Redistricting
date: last-modified
authors:
  - name: Cory McCartan
    affiliations:
      - name: Pennsylvania State University
        department: Department of Statistics
        address: 461 Pollock Road
        city: University Park
        region: PA
        postal-code: 16801
    email: mccartan@psu.edu
    url: https://corymccartan.com/
    corresponding: true
  - name: Christopher T. Kenny
    affiliations:
    - name: Harvard University
      department: Department of Government
      address: 1737 Cambridge Street
      city: Cambridge
      region: MA
      postal-code: "02138 "
    email: christopherkenny@fas.harvard.edu
    url: https://christophertkenny.com
abstract: |
  Social scientists have developed dozens of measures for assessing partisan bias in redistricting.
  But these measures are not easily adapted to other groups, 
  including groups defined by race, class, or geography.
  Nor are they applicable to single- or no-party contexts, such as local redistricting.
  To overcome these limitations, we propose a unified framework of harm for evaluating 
  the impacts of a districting plan on individual voters and the groups to which they belong. 
  We consider a voter harmed if their chosen candidate is not elected under the current plan, 
  but would be under a different plan.
  Harm improves on existing measures by both focusing on the choices of individual voters 
  and directly incorporating counterfactual plans.
  We discuss strategies for estimating harm, and demonstrate the utility of our framework 
  through analyses of partisan gerrymandering in New Jersey, voting rights 
  litigation in Alabama, and racial dynamics of Boston City Council elections.
thanks: |
  The authors thank Steve Ansolabehere, Matt Blackwell, Kosuke Imai, Gary King, 
  Shiro Kuriwaki, Tyler Simko, Nick Stephanopolous, and 
  participants of the Fall 2021 Harvard American Politics Research Workshop,
  Fall 2022 Harvard Applied Statistics Workshop, 
  and Imai Research Group for helpful comments and discussion.
  Replication code and data are available at <https://github.com/CoryMcCartan/harm-redistricting>.
keywords:
  - representation
  - redistricting
  - voting rights
  - individual harm
bibliography: references.bib
format: 
    cmc-article-pdf:
        fontsize: 11pt
        biblio-style: apsr
        include-in-header: "header.tex"
editor_options:
  markdown:
    wrap: sentence
---

@@

>>>>

---
title: Individual and Differential Harm in Redistricting
date: last-modified
authors:
  - name: Cory McCartan
    affiliations:
      - name: Pennsylvania State University
        department: Department of Statistics
        address: 461 Pollock Road
        city: University Park
        region: PA
        postal-code: 16801
    email: mccartan@psu.edu
    url: https://corymccartan.com/
    corresponding: true
  - name: Christopher T. Kenny
    affiliations:
    - name: Harvard University
      department: Department of Government
      address: 1737 Cambridge Street
      city: Cambridge
      region: MA
      postal-code: "02138 "
    email: christopherkenny@fas.harvard.edu
    url: https://christophertkenny.com
abstract: |
  Social scientists have developed dozens of measures for assessing partisan bias in redistricting.
  But these measures are not easily adapted to other groups, 
  including groups defined by race, class, or geography.
  Nor are they applicable to single- or no-party contexts, such as local redistricting.
  To overcome these limitations, we propose a unified framework of harm for evaluating 
  the impacts of a districting plan on individual voters and the groups to which they belong. 
  We consider a voter harmed if their chosen candidate is not elected under the current plan, 
  but would be under a different plan.
  Harm improves on existing measures by both focusing on the choices of individual voters 
  and directly incorporating counterfactual plans.
  We discuss strategies for estimating harm, and demonstrate the utility of our framework 
  through analyses of partisan gerrymandering in New Jersey, voting rights 
  litigation in Alabama, and racial dynamics of Boston City Council elections.
thanks: |
  The authors thank Steve Ansolabehere, Matt Blackwell, Kosuke Imai, Gary King, 
  Shiro Kuriwaki, Tyler Simko, Nick Stephanopolous, and 
  participants of the Fall 2021 Harvard American Politics Research Workshop,
  Fall 2022 Harvard Applied Statistics Workshop, 
  and Imai Research Group for helpful comments and discussion.
  Replication code and data are available at <https://github.com/CoryMcCartan/harm-redistricting>.
keywords:
  - representation
  - redistricting
  - voting rights
  - individual harm
bibliography: references.bib
format: 
    cmc-article-pdf:
        journal:
            blinded: true
        fontsize: 11pt
        biblio-style: apsr
        include-in-header: "header.tex"
editor_options:
  markdown:
    wrap: sentence
---

@@

# Data availability {.unnumbered}
All data are publicly available and are included in the replication package for the manuscript.

# Code availability {.unnumbered}
All code is included in the replication package for the manuscript.
